Unrecognized States: A Theory of Self-Determination and Foreign Influence

### 2015-09-08 draft.tex
fix lemmas 3 and 4 for alpha and nu positive

\item  \textit{$-\nu\cdot p_{1s} \geq \lambda \cdot p_{1s} + \mu + \beta$: recognition of the secessionist state is more important for the international community to avoid than for the patron to achieve.}

@@

Given our assumption that the status quo outcome is no worse than the reunification outcome net of investments for player $s$, when $\frac{\alpha}{1 -\de} > \frac{\beta+ R_c^{SQ}}{1 -\de}$, $p$ is willing to invest enough to create a large enough difference in $Q_{sn} - l_{sn}$ so that player $c$ will not find contesting the status quo to be in its interest. Whenever the status quo will be the outcome in this way, $c$ maximizes its utility by choice of $R_c^{SQ} =0$. Thus a sufficient condition to prevent $c$ from inducing the \emph{reunification} outcome is $\alpha  > \beta$. \hfill $\blacksquare$
\\
\\
\noindent {\bf Proof of Lemma 3} \\
Again, in the status quo, we have $U_{pn}^{SQ}= -R_p^{SQ}$ and $U_{cn}^{SQ}= -R_c^{SQ}$. If the outcome is recognition of the secessionists, per-period payoffs in this absorbing state become $U_{pn}^{REC}= \lambda -R_p^{REC}$ and $U_{cn}^{REC}= \nu -R_c^{REC}$ where $SQ$ and $REC$ distinguish investments under the status quo and recognition scenarios respectively.

Thus the difference in continuation values from switching from the status quo to recognition for player $p$ is $\frac{\lambda +R_p^{SQ}}{1 -\de} -R_p^{REC}$. For player $c$ it is $\frac{\nu + R_c^{SQ}}{1 -\de}-R_c^{REC}$. As noted above, $R_c^{REC} =0$. So $c$ is willing to invest up to $-\nu$ per period toward maintaining the status quo, and $p$ is willing to invest up to $\frac{\lambda +R_p^{SQ}}{1 -\de}$ toward recognition. We have shown that the largest possible investments by $p$ in the status quo are $\frac{\beta}{1 -\de}$ in $n=1$ and $\mu$ in each period thereafter, so the largest this difference can be for $p$ is $\frac{\lambda + \mu + \beta}{1 -\de}$.

Player $c$ would be willing to counter whatever investment player $p$ makes to try to induce recognition---e.g. counter an addition to $l_{gn}$ with an investment to $q_{gn}$---as long as $-\nu \geq L_{gn} - q_{gn}$, where $L_{gn}$ is gross of any investment by $p$.\footnote{As in the proof of Lemma 2, we have assumed that the payoff relationship at $n$ is such that ceding is better than fighting. If this relationship is reversed, similar analysis and results hold.} That is, the required investment must be no more than the loss player $c$ receives from investing.

@@

Thus $\frac{-\nu}{1 -\de} > \frac{\lambda + \mu + \beta}{1 -\de}$ guarantees that player $c$ will always be willing to counter an investment by player $p$ toward achieving recognition, making $p$ unwilling to make that investment in the initial stage and making the response by player $c$ unnecessary. A slightly weaker version of this inequality is stated as condition 4 so that this deterrence dynamic is guaranteed to operate through the war lottery as well, as will be seen in Lemma 4. \hfill $\blacksquare$
\\
\\
\noindent {\bf Proof of Lemma 4} \\
The difference in continuation values from switching from the status quo to taking the war lottery for player $p$ is $\frac{\lambda p_{1s} + \alpha (1-p_{1s})}{1 -\de} +R_p^{SQ} -R_p^{WAR}$. For player $c$ it is $\frac{\nu p_{1s} + \beta (1-p_{1s})}{1 -\de} + R_c^{SQ} -R_c^{WAR}$. As noted above, $R_c^{SQ} =0$. So $c$ is willing to invest up to $\frac{\nu p_{1s} + \beta (1-p_{1s})}{1 -\de}$ toward preventing fighting if this quantity is negative (leave the case where it is non-negative until later).

$p$ is willing to invest up to $\frac{\lambda p_{1s} + \alpha (1-p_{1s})}{1 -\de} +R_p^{SQ}$ toward instigating fighting when this quantity is positive, and will not invest when the quantity is non-positive. Thus we must only examine the case where this quantity is positive.

We will again set $R_p^{SQ}$ at its largest value of $\frac{\mu + \beta}{1-\de}$ as this is most restrictive. In this only case of interest, we have $\frac{\alpha (1-p_{1s}) + \lambda p_{1s} + \mu + \beta}{1 - \de}> 0$. Note that the first term is negative while the rest are positive.

Multiplying Condition 3 by $-(1-p_{1s})$, we have $-\beta \cdot (1-p_{1s}) \geq \alpha \cdot (1-p_{1s})$. Using this with Condition 4, we have $- \nu p_{1s} - \beta (1-p_{1s}) \geq \alpha (1-p_{1s}) +\lambda p_{1s} + \mu + \beta> 0 \ \forall p_{1s},p_{2s}$ and therefore also
\[
  \textstyle -\left(\frac{\nu p_{1s} + \beta (1-p_{1s})}{1 -\de}\right) \geq \frac{\alpha (1-p_{1s})+ \lambda p_{1s} + \mu + \beta}{1 - \de}> 0 \ \forall p_{1s},p_{2s}
\]

>>>>

\item  \textit{$\nu\cdot p_{1s} \geq \lambda \cdot p_{1s} + \mu + \beta$: recognition of the secessionist state is more important for the international community to avoid than for the patron to achieve.}

@@

Given our assumption that the status quo outcome is no worse than the reunification outcome net of investments for player $s$, when $\frac{\alpha}{1 -\de} > \frac{\beta+ R_c^{SQ}}{1 -\de}$, $p$ is willing to invest enough to create a large enough difference in $Q_{sn} - l_{sn}$ so that player $c$ will not find contesting the status quo to be in its interest. Whenever the status quo will be the outcome in this way, $c$ maximizes its utility by choice of $R_c^{SQ} =0$. Thus a sufficient condition to prevent $c$ from inducing the \emph{reunification} outcome is $\alpha  > \beta$. \hfill $\blacksquare$
\\
\\
\noindent {\bf Proof of Lemma 3} \\
Again, in the status quo, we have $U_{pn}^{SQ}= -R_p^{SQ}$ and $U_{cn}^{SQ}= -R_c^{SQ}$. If the outcome is recognition of the secessionists, per-period payoffs in this absorbing state become $U_{pn}^{REC}= \lambda -R_p^{REC}$ and $U_{cn}^{REC}= -\nu -R_c^{REC}$ where $SQ$ and $REC$ distinguish investments under the status quo and recognition scenarios respectively.

Thus the difference in continuation values from switching from the status quo to recognition for player $p$ is $\frac{\lambda +R_p^{SQ}}{1 -\de} -R_p^{REC}$. For player $c$ it is $\frac{-\nu + R_c^{SQ}}{1 -\de}-R_c^{REC}$. $R_c^{REC} =0$ since the international community will not invest to create its least preferred outcome. So $c$ is willing to invest up to $\nu$ per period toward maintaining the status quo, and $p$ is willing to invest up to $\frac{\lambda +R_p^{SQ}}{1 -\de}$ toward recognition. We have shown that the largest possible investments by $p$ in the status quo are $\frac{\beta}{1 -\de}$ in $n=1$ and $\mu$ in each period thereafter, so the largest this difference can be for $p$ is $\frac{\lambda + \mu + \beta}{1 -\de}$.

Player $c$ would be willing to counter whatever investment player $p$ makes to try to induce recognition---e.g. counter an addition to $l_{gn}$ with an investment to $q_{gn}$---as long as $\nu \geq L_{gn} - q_{gn}$, where $L_{gn}$ is gross of any investment by $p$.\footnote{As in the proof of Lemma 2, we have assumed that the payoff relationship at $n$ is such that ceding is better than fighting. If this relationship is reversed, similar analysis and results hold.} That is, the required investment must be no more than the loss player $c$ receives from investing.

@@

Thus $\frac{\nu}{1 -\de} > \frac{\lambda + \mu + \beta}{1 -\de}$ guarantees that player $c$ will always be willing to counter an investment by player $p$ toward achieving recognition, making $p$ unwilling to make that investment in the initial stage and making the response by player $c$ unnecessary. A slightly weaker version of this inequality is stated as condition 4 so that this deterrence dynamic is guaranteed to operate through the war lottery as well, as will be seen in Lemma 4. \hfill $\blacksquare$
\\
\\
\noindent {\bf Proof of Lemma 4} \\
The difference in continuation values from switching from the status quo to taking the war lottery for player $p$ is $\frac{\lambda p_{1s} - \alpha (1-p_{1s})}{1 -\de} +R_p^{SQ} -R_p^{WAR}$. For player $c$ it is $\frac{-\nu p_{1s} + \beta (1-p_{1s})}{1 -\de} + R_c^{SQ} -R_c^{WAR}$. As noted above, $R_c^{SQ} =0$. So $c$ is willing to invest up to $\frac{-\nu p_{1s} + \beta (1-p_{1s})}{1 -\de}$ toward preventing fighting if this quantity is negative (leave the case where it is non-negative until later).

$p$ is willing to invest up to $\frac{\lambda p_{1s} - \alpha (1-p_{1s})}{1 -\de} +R_p^{SQ}$ toward instigating fighting when this quantity is positive, and will not invest when the quantity is non-positive. Thus we must only examine the case where this quantity is positive.

We will again set $R_p^{SQ}$ at its largest value of $\frac{\mu + \beta}{1-\de}$ as this is most restrictive. In this only case of interest, we have $\frac{\lambda p_{1s} - \alpha (1-p_{1s}) + \mu + \beta}{1 - \de}> 0$.

Multiplying Condition 3 by $-(1-p_{1s})$, we have $-\beta \cdot (1-p_{1s}) \geq -\alpha \cdot (1-p_{1s})$. Using this with Condition 4, we have $\nu p_{1s} - \beta (1-p_{1s}) \geq \lambda p_{1s} - \alpha (1-p_{1s}) + \mu + \beta> 0 \ \forall p_{1s},p_{2s}$ and therefore also
\[
  \textstyle -\left(\frac{-\nu p_{1s} + \beta (1-p_{1s})}{1 -\de}\right) \geq \frac{\lambda p_{1s} - \alpha (1-p_{1s})+ \mu + \beta}{1 - \de}> 0 \ \forall p_{1s},p_{2s}
\]
